\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to express my profound gratitude to a number of people whose support and contributions were invaluable throughout the course of this thesis.

First and foremost, I would like to thank my supervisor and mentor, Dr. Haris Volos, for his unwavering guidance and continuous support. Whenever issues arose, Dr. Volos was always ready to thoroughly evaluate the problem, imparting not only his expert knowledge but also a sense of shared purpose and curiosity. His dedication to my research gave me the confidence and motivation to navigate the complex challenges that arose during this thesis. It truly felt like we shared the same research goals and the drive to satisfy curiosity which is crucial in a research context. His care and attention were instrumental in enabling me to reach my research goals.

I am deeply grateful to my university for granting me access to the CloudLab's infrastructure. This access was not just a privilege but a necessity. The comprehensive results and the eventual completion of this thesis would not have been possible without it. It provided a robust platform to explore Swift as a serverless language on OpenWhisk, contributing significantly to the outcomes of my research.

Lastly, but by no means least, I would like to extend my heartfelt thanks to my parents. Their unconditional faith in my abilities and their financial support provided a stable foundation for my academic journey. It allowed me to comfortably pursue my interests and devote my energy to my research. Their sacrifices and belief in my vision were a beacon of light that guided me through the most challenging phases of this thesis.

To all of you, thank you. This accomplishment would not have been possible without your collective support and belief.

